Group polarization replications may be marred by high false discovery rates

### 2025-01-01 GroupPolarizationStatmod/paper/ms.tex
nearly done with Analysis

To calculate the false detection rate we begin by simulating 
pre- and post-deliberation 
opinions by drawing $N_e$ samples from normal latent opinion
distributions that generated spurious group polarization, i.e., with
parameters labelled $\beta_e$ found in the hillclimbing step. Binning these
random values simulates observations. $N_e$ is the number
of participants in experimental condition $e$. Each trial $i$ of $N_e$
simulated pre- and post-deliberation observations is then inspected to see
if spurious group polarization occurred in the simulation. Spurious group
polarization is determined by first inferring the latent parameters of each
simulated experimental trial by fitting an ordered probit model~\cite{Liddell2018}. 
The inferred parameters are then used to calculate the effect size for that
trial, Cohen's $d$, defined as~\cite[p. 331]{Liddell2018}
\begin{equation}
  d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
  \label{eq:cohens}
\end{equation}
\noindent
The family-wise error rate is, $\alpha_{e} = \Pr(D|\neg E)$, is the 
fraction of trials, 
with $d_{ei} > d^*$ for the $i^\text{th}$ trial for experimental
condition $e$, written
\begin{equation}
  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta_i,
\end{equation}
\noindent
where
\[
  \Theta_i = 
    \begin{cases}
      1 & \text{ if } d_{ei} > d^* \\
      0 & \text{ otherwise.}
    \end{cases}
\]
and $N_T = 1000$ is the number of trials.

To calculate the family-wise error rate
for a study, $s$ which contains several experimental conditions, $e$,
\begin{equation}
  \alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e
\end{equation}
\noindent
The grand mean across all studies for the corpus is 
\begin{equation}
  \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
\end{equation}
With $\alpha$ calculated, we calculate the false discovery rate using
Equation~\ref{eq:fdr}.

We invert this procedure to find what significance value $d^*$ is necessary
to achieve a low family-wise error rate $\alpha_e = 0.05$ for experimental
condition, $e$. To do this we first calculate $\alpha_e(d^*)$ for a
sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
$\alpha_e(d^*) \leq 0.05$.

@@

Our analysis estimates how the probability that a published detection of group
polarization is actually a false detection. We first identified latent opinion
parameters that yield spurious group polarization in an asymptotic $N\to\infty$
model. This gives us our first suggestion that experimental conditions may be
plausibly false. Briefly, we identified latent opinion distribution parameters
that generated spurious group polarization in 54 out of 57 experimental
conditions. In other words, 95\% of all group polarization detections are
plausibly spurious due to ceiling artifacts from the continuous latent
opinions-ordinal measurement experimental design. We then analyze the family-wise
error ($\alpha$) and false detection rate (FDR) 
for each experimental condition and aggregated 
up to the study level and across all studies and experimental conditions. 
The lower twentieth percentile of $\alpha$
We close by showing that $d^*$ must be greater than the ``high'' effect size
$d^*=0.8$ to limit family-wise error rates to $\alpha \leq 0.05$ in all but six
experimental conditions.

\subsection{Nearly all group polarization detections are plausibly false}

We used our formal distribution model of group polarization experiments
to identify plausible latent simple consensus distributions that generate spurious
group polarization in observed ordinal distributions 
in the asymptotic case of taking $N \to \infty$ ordinal measurements.
Across the ten studies we analyzed, we excluded 3 out of 60 experimental
conditions that did not claim to be positive findings of group polarization.
Of the remaining 57, we identified plausible $\beta$ that generate spurious
group polarization in 54 experimental conditions. In the other three, we did
identify $\beta$ that generate spurious group polarization, but the initial
simulated distributions were highly polarized. Two of these
simulations of experimental conditions from Schkade, et al., (2010) because
the histogram plots of the observed opinion distributions are not
polarized~\cite[Figure 1, p. 234]{Schkade2010}. The other condition where we
rejected candidate $\beta$ was in a question about a ``bad teacher'' in Myers
(1975)~\cite{Myers1975a}, where lower values signify increasingly 
worse impressions of the teacher and higher values signify increasingly good
impressions. It is unlikely that any participants would rate a ``bad teacher''
as good, so we excluded this from our analysis as well. This exclusion may be
overly generous since Myers (1975) did not provide details beyond mean values
and opinion shifts, and the article was published before the importance
of open data best practices was recognized.

\begin{table}[ht]
  \caption{yo}
  \label{tab:quantiles}
\rotatebox{90} {\textbf{Quantiles for the two error-rate measures under different
  significance values.}\mt{Not sure about the best way to format; probably don't
  need this in the main text, but could be valuable in supplement?}}
  \input{tex/quantile_table.tex}

\end{table}

\subsection{Half of experiments exceed a 75\% false discovery rate at ``medium'' significance}

We simulated trials from each experimental condition, $e$, 
by sampling and binning pre- and
post-deliberation latent opinions from distributions with parameters identified
with the $N\to\infty$ model as causing plausibly spurious group
polarization. For each $e$,
we drew a number of samples equal to the number of participants in 
the original experiments, $N_\mathrm{emp}$. To estimate
family-wise error and false detection rates, we fit an ordered probit
model to the simulated observations and obtained estimates of the group polarization
effect size, $d$ (Equation~\ref{eq:cohens}). If the effect size was greater than a
significance value, $d > d^*$, then by definition there was a detection $D$ of group 
polarization,
even though there was no effect by design, written $\neg E$. 
The family-wise error rate is $\Pr(D|\neg E)$, calculated as the frequency with which $d > d^*$ out of 1000
simulation trials. The false detection rate is the family-wise error rate scaled
using the base rate of effects and the statistical power (Equation~\ref{eq:fdr}). We
test three difference significance values, $d = 0.2,~0.5$ and $0.8$, corresponding
to what Cohen prescribed to use for ``low'', ``medium'', and ``high'' significance.

Family-wise error rates (Figure~\ref{fig:fwer_fdr_synthesis} and
Table~\ref{tab:quantiles}) varied from a minimum of one in ten for the
\texttt{Feminists-Experimental} condition from Myers (1975)~\cite{Myers1975a}
to a maximum of two in three for the \texttt{COSprings-CivilUnions} condition from
Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010}. The median family-wise error
rate is 0.28 for Krizan and Baron's (2007) experimental condition
\texttt{NoOutgroupScenario1}~\cite{Krizan2007}.

@@

  \centering
  \includegraphics[width=1.1\textwidth]{Figures/Results/fwer_fdr_synthesis.pdf}

\end{figure}

@@

Solving the inverse problem of which $d^*$ achieves a low family-wise error rate,
which drives the false detection rate. Here we present our calculations to find
$d^*$ that achieve $\alpha(d^*) = 0.05$, a commonly-assumed parameter value of
metascience.

\begin{figure}
  \caption{\textbf{Significance value, $d^*$, to limit family-wise error 
  rate of 0.05 for each identified study.} \mt{Does this make sense? Order by the
rankings in fdr\_vs\_experiment plots to better understand/compare.}}
  \centering
    \includegraphics[width=0.75\textwidth]{Figures/Results/sigval_for_low_fwer.pdf}
  \label{fig:sigval_for_low_fwer}
\end{figure}

@@

% \section{Figures}

% \begin{figure}
%   \caption{
%     Boxplots visualizing the distribution of Cohen's $d$ values (x-axis in
%     A and B) inferred through a metric linear model over 1000 simulation 
%     trials for each experimental condition (y-axis in A and B). All conditions
%     have many trials that result in non-zero Cohen's $d$ values with number of
%     simulated participants, $N$, set to their empirical values reported in the
%     associated journal article (A). However, when the number of simulated
%     participants is increased to $10N$, there are several conditions that 
%     indeed yield distributions closer to $d = 0$ as expected, however some
%     conditions have $d$ distributions centered on a nonzero mean with little to
%     no overlap with $d=0$; variance in $d$ decreases as expected
%     across experimental conditions (B).
%   }  
%   \label{fig:MetricBoxplots}
%   \centering
%   \begin{subfigure}[]{0.9\textwidth}
%     % \hspace{-7.5em} 
%     \centering
%     \textsf{\large \textbf{A}, empirical $N$} \\
%     \includegraphics[width=\textwidth]{Figures/boxplots/metric_cohens.pdf}
%     \label{fig:MetricBoxplot_N}
%   \end{subfigure} \\
% \end{figure}
% \begin{figure}\ContinuedFloat
%   \begin{subfigure}[]{0.9\textwidth}
%     % \hspace{-5.5em} 
%     \centering
%     \textsf{\large \textbf{B}, 10 $\times$ empirical $N$} \\
%     \includegraphics[width=\textwidth]{Figures/boxplots/10N_metric_cohens.pdf}
%     \label{fig:MetricBoxplot_10N}
%   \end{subfigure}
% \end{figure}

% \begin{figure}
%   \caption{Boxplot of Cohen's $d$ for the ordered probit model over 100
%     simulation trias for each experimental condition (y-axis in A and B). Each
%     trial represents a possible outcome of a group polarization experiment where
%     the true opinion shift is zero. The mean across trials is closer to 0 than
%     in the metric case, but many simulated zero-shift experiments still result in
%     an observed shift. This illustrates the weak statistical power of these group
%     polarization experiments.}
%   \label{fig:OrdinalBoxplot}
%   \centering
%     \includegraphics[width=0.9\textwidth]{Figures/boxplots/ordinal_cohens.pdf}
% \end{figure}

>>>>

To calculate the false detection rate we begin by simulating 
pre- and post-deliberation 
opinions by drawing $N_e$ samples from normal latent opinion
distributions that generated spurious group polarization, i.e., with
parameters labelled $\beta_e$ found in the hillclimbing step. Binning these
random values simulates observations. $N_e$ is the number
of participants in experimental condition $e$. Each trial $i$ of $N_e$
simulated pre- and post-deliberation observations is then inspected to see
if spurious group polarization occurred in the simulation. Spurious group
polarization is determined by first inferring the latent parameters of each
simulated experimental trial by fitting an ordered probit model~\cite{Liddell2018}. 
The inferred parameters are then used to calculate the effect size for that
trial, Cohen's $d$, defined as~\cite[p. 331]{Liddell2018}
\begin{equation}
  d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
  \label{eq:cohens}
\end{equation}
\noindent
The family-wise error rate is, $\alpha_{e} = \Pr(D|\neg E)$, is the 
fraction of trials, 
with $d_{ei} > d^*$ for the $i^\text{th}$ trial for experimental
condition $e$, written
\begin{equation}
  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
  \label{eq:study_aggregate}
\end{equation}
\noindent
where
\[
  \Theta(d_{ei}; d^*) = 
    \begin{cases}
      1 & \text{ if } d_{ei} > d^* \\
      0 & \text{ otherwise.}
    \end{cases}
\]
and $N_T = 1000$ is the number of trials. For simplicity we assume that the sign of
$d_{ei}$ is positive, meaning a shift to greater extremisim. It is possible that
$d_{ei}$ could be negative, indicating a shift to lesser extremity.

To calculate the family-wise error rate
for a study, $s$ which contains several experimental conditions, $e$,
\begin{equation}
  \alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e
\end{equation}
\noindent
The mean across all results identified as plausibly spurious is 
\begin{equation}
  \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
\end{equation}
We finally may calculate the false discovery rate by plugging this $\alpha$
into Equation~\ref{eq:fdr} and assuming empirically-motivated values for 
base rate, $b$, and statistical power, $W$, as well as more pessimistic and
optimistic settings for base rate and power.

We close our Analysis by inverting this procedure to find what significance value
$d^*$ is necessary to achieve a low family-wise error rate $\alpha_e = 0.05$ for
experimental condition, $e$. To do this we first calculate $\alpha_e(d^*)$ for a
sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
$\alpha_e(d^*) \leq 0.05$.

@@

Our analysis estimates how the probability that a published detection of group
polarization is actually a false detection. We first identified latent opinion
parameters that yield spurious group polarization in an asymptotic $N\to\infty$
model. This gives us our first suggestion that experimental conditions may be
plausibly false. Briefly, we identified latent opinion distribution parameters
that generated spurious group polarization in 54 out of 57 experimental
conditions. In other words, 95\% of all group polarization detections are
plausibly spurious due to ceiling artifacts from the continuous latent
opinions-ordinal measurement experimental design. We then analyze the family-wise
error ($\alpha$) and false detection rate (FDR) 
for each experimental condition and aggregated 
up to the study level and across all studies and experimental conditions. 
We estimate the lower twentieth percentile of experimental conditions 
have \emph{at least} $\alpha \leq 0.16$, 
with median $\alpha \geq 0.27$ and upper eightieth percentile with $\alpha \geq
0.40$ (Table~\ref{tab:quantiles}\mt{A.  –but don't have A and B; TODO: separate
FWER and FDR tables, only use 0, 20, 50, 80, 100 for probs. Rename 0 -> min, 100 ->
max.}). For the fiftieth and eightieth percentiles this
corresponds to false discovery rates of 0.75 and 0.82, respectively. 
We emphasize that these estimates and others detailed below are lower bounds on
$\alpha$ and FDR since we are using what is likely an overgeneralized,
underspecified statistical model that does not account for all the potential
variance in experimental outcomes, which would increase error rates.
We close our Analysis by calculating which $d^*$ limits family-wise error rates to
$\alpha \leq 0.05$. This is greater than $d^* = 0.8$, classically taken to 
represent ``high'' significance, in all but six experimental conditions.

\subsection{Nearly all group polarization detections are plausibly spurious}

We used our formal distribution model of group polarization experiments
to identify plausible latent simple consensus distributions that generate spurious
group polarization in observed ordinal distributions 
in the asymptotic case of taking $N \to \infty$ ordinal measurements.
Across the ten studies we analyzed, we excluded 3 out of 60 experimental
conditions that did not claim to be positive findings of group polarization.
Of the remaining 57, we identified simple consensus $\beta$ that generate spurious
group polarization in 54 experimental conditions. In the other three, we did
identify $\beta$ that generate spurious group polarization, but the initial
simulated distributions were highly polarized. We rejected $\beta$ generated for
two spurious group polarization from Schkade, et al., (2010) because
the histogram plots of the observed opinion distributions are not
polarized~\cite[Figure 1, p. 234]{Schkade2010}. The other condition where we
found but rejected spurious group polarization $\beta$ was in a question 
about a ``bad teacher'' in Myers (1975)~\cite{Myers1975a}, where lower values
signify increasingly worse impressions of the teacher and higher values signify
increasingly good impressions. It is unlikely that a
``bad teacher'' would be rated as good in this case, so we excluded this 
from our analysis as well. This exclusion may be overly generous since Myers (1975)
did not provide details beyond mean values and opinion shifts, and the article was
published before open data practices.

\begin{table}[ht]
  \caption{{\textbf{Quantiles for the two error-rate measures under different
  significance values.}\mt{Working on splitting in two and having more base rate
and power conditions.}}}
  \label{tab:quantiles}
\rotatebox{90} {
  \input{tex/quantile_table.tex}
}
\end{table}

\subsection{Half of experiments exceed a 70\% false discovery rate at ``medium'' significance}

We simulated trials from each experimental condition, $e$, by sampling and binning
pre- and post-deliberation latent opinions from distributions with parameters
identified with the $N\to\infty$ model as causing plausibly spurious group
polarization. For each $e$, we drew a number of samples equal to the number of
participants in the original experiments, $N_\mathrm{emp}$. To estimate family-wise
error and false detection rates, we fit an ordered probit model to the simulated
observations and obtained estimates of the group polarization effect size, $d$
(Equation~\ref{eq:cohens}). If the effect size was greater than a significance
value, $d > d^*$, then by definition there was a detection $D$ of group
polarization, even though there was no effect by design, written $\neg E$.  The
family-wise error rate is $\Pr(D|\neg E)$, calculated as the frequency with which $d
> d^*$ out of 1000 simulation trials. The false detection rate is the family-wise
error rate scaled using the base rate of effects and the statistical power
(Equation~\ref{eq:fdr}). We test three difference significance values, $d =
0.2,~0.5$ and $0.8$, corresponding to what Cohen prescribed to use for ``low'',
``medium'', and ``high'' significance. If not otherwise noted, we set the base
rate to be $b=0.1$ and the power $W = 0.8$ as metascience parameters to calculate 
the false discovery rate (Equation~\ref{eq:fdr}). Note that if the family-wise error
rate is low at $\alpha = 0.05$, the false discovery rate is $\fdr = 0.36$.

Different conditions yield different effect size distributions
(Figure~\ref{fig:effect_size_distros}), which 
determine the magnitude of error rates and difference between the error rates for low,
medium, and high significance values $d^*=0.2,0.5,0.8$ magnitude of error rates 
(Figure~\ref{fig:fwer_fdr_synthesis}B). 
Family-wise error rates varied from a minimum of one in ten for the
\texttt{Feminists-Experimental} condition from Myers (1975)~\cite{Myers1975a}
to a maximum of two in three for the \texttt{COSprings-CivilUnions} condition from
Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010} under medium
significance effect size $d^* = 0.5$ (Figure~\ref{fig:fwer_fdr_synthesis} and
Table~\ref{tab:quantiles}). The median family-wise error
rate is 0.28 for Krizan and Baron's (2007) experimental condition
\texttt{NoOutgroupScenario1}~\cite{Krizan2007}. This translated to a minimum
false detection rate of 0.52 for the Myers (1975) condition, a maximum false
detection rate of 0.88 for the Schkade, et al., (2010) condition, and a median false
detection rate of 0.75 for the Krizan and Baron (2007) condition.

In the conditions
with lower family-wise error rates, there were several conditions where using the low 
significance value $\alpha$ resulted in a much greater error rate than the medium or 
high significance values. In the ten conditions with the lowest $\alpha$ for 
$d^* = 0.5$, with $\alpha$ near 0.1, five of these conditions have $\alpha > 0.25$
when $d^* = 0.2$ corresponding to a small effect size. For low $\alpha$, the
difference in false detection rates for each $d^*$ value are more widely
distributed, following the distribution of $\alpha$, meaning that 
a modest increase in $d^*$ can have greater effects on the false discovery rate. In
the conditions above the 50th percentile, $\alpha$ approaches and exceeds one in
two, even when requiring ``high'' significance with $d^*=0.8$. With $d^* = 0.8$,
$\alpha$ approaches and exceeds one in four, with four near one in two. When 
requiring ``low'' signifance,

Experimental designs within each study or article tend to share certain
characteristics, such as the number of bins in the ordinal measurement design
or the number of participants per condition. Therefore, the 
average false detection rate across experimental trials within a study
(Equation~\ref{eq:study_aggregate}) show that

@@

  \centering
  \includegraphics[width=1.1\textwidth]{Figures/Analysis/fwer_fdr_synthesis.pdf}
\end{figure}

\begin{itemize}
  \item 
    Other $b$ and $W$ values in quantiles.
  \item
    Note that there were many ``significant'' results---in the wrong direction
    (Figure~\ref{fig:OrdinalBoxplot}).
\end{itemize}

@@

Solving the inverse problem of which $d^*$ achieves a low family-wise error rate,
which drives the false detection rate. Here we present our calculations to find
$d^*$ that achieve $\alpha(d^*) = 0.05$.

\begin{figure}
  \caption{
    \textbf{Significance value, $d^*$, to limit family-wise error 
    rate of 0.05 for each identified study.} The Moscovici and Zavalloni (1969)
    ``Americans'' question would require the greatest $d^*$ to achieve $\alpha \leq
    0.05$, even though it does not have the highest error rates. This is because of
    larger outliers in simulated $d$ for this condition compared to others with
    greater error rates (Figure~\ref{fig:OrdinalBoxplot}).
  }
  \centering
    \includegraphics[width=0.75\textwidth]{Figures/Analysis/sigval_for_low_fwer.pdf}
  \label{fig:sigval_for_low_fwer}
\end{figure}

@@

\renewcommand{\thefigure}{A\arabic{figure}}

\setcounter{figure}{0}

\clearpage

%%%%%%%%%% Supplement %%%%%%%%%%
\pagebreak
\begin{center}
  \textbf{\Large \textsf{Minority-group incubators and majority-group
reservoirs support the diffusion of climate change adaptations (Supplementary Material)}}
\end{center}

%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
\setcounter{equation}{1}
\setcounter{figure}{0}
\setcounter{section}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\makeatletter
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thepage}{S\arabic{page}}

\section{Analysis}

\begin{figure}
  \caption{Boxplot of Cohen's $d$ for the ordered probit model over 1000
    simulation trias for each experimental condition (y-axis in A and B). Each
    trial represents a possible outcome of a group polarization experiment where
    the true opinion shift is zero. The mean across trials is closer to 0 than
    in the metric case, but many simulated zero-shift experiments still result in
    an observed shift. This illustrates the weak statistical power of these group
    polarization experiments.}
  \label{fig:OrdinalBoxplot}
  \centering
    \includegraphics[width=1.0\textwidth]{Figures/Analysis/ordinal_cohens.pdf}
\end{figure}
